---
title: "week16_recap"
author: "Daniel Trielli"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(tigris)
library(tidycensus)
library(corrr)
library(sf) 
library(tidytext)
```

## Load Data

[USAFacts compiled data on inspections on child care facilities throughout the country](https://usafacts.org/data-projects/childcare-inspections) - specifically, how many of those facilities are overdue for inspection. Let's take a look on the data they have on Maryland.

First, here the data we need to load:

```{r}
childcare_inspections <- read_csv('data/childcare_inspections.csv')
childcare_inspections$fips <- as.character(childcare_inspections$fips)
```

#### **Q1** The childcare data does not have the names of the counties. Join them with the md_counties data, that can be loaded in the block below. Then inspect the dataframe you created with this join and answer this: there are two columns with county names in the new dataframe. Which one is the best to be used? (hint: it has to do with homonyms)

```{r}
# This loads the county data
md_counties <- counties() |>  filter(STATEFP == "24")
```

```{r}
# Join the childcare data with the county data 
maryland_childcare_counties <- md_counties |> 
  inner_join(childcare_inspections, join_by(GEOID == fips))
```

##### A1: Nameslad because it differentiates Baltimore County and Baltimore City

------------------------------------------------------------------------

#### **Q2** Create bar chart with the percentages of overdue childcare facility inspections per county to answer (You'll need to create a column for that calculation first). What is the county with the largest percentage of facilities with overdue inspections?

```{r}
# Calculate the percentage of overdue inspections per county
maryland_childcare_counties <- maryland_childcare_counties |>
  mutate(pct = (overdue/total) *100) 
# Create the chart with ggplot
maryland_childcare_counties |>
  ggplot() + 
  geom_bar(aes(x=reorder(NAMELSAD,pct), weight=pct)) + 
  coord_flip() +
  labs(
    title="Maryland Counties with the most overdue childcare facility inspections",
    y = "# of overdue facilities",
    x = "county",
  )
  

```

##### **A2: Montgomery county has the largest percentage of overdue facilities**

------------------------------------------------------------------------

#### **Q3** Next, we want to see if the number of child care facilities - with overdue inspections or otherwise - is appropriate for each county. So we're going to calculate the rate of children per facility per county. To do that, you'll first need Census data to get the population of children under 5 for each county. Which county has the biggest rate of children per facility (i.e. they could use more facilities)?

```{r}
# First, grab data from the ACS. The ACS variable for children age under 5 is B06001_002 We want 2022 data.
  md <- get_acs(geography = "county",
              variables = c(children_under_5 = "B06001_002"),
              state = "MD",
              year = 2022)
```

```{r}
# Rename the column name 'estimate', which has our population data, with something more useful, like 'children_pop'
md |>
  rename(children_pop = estimate)
```

```{r}
# Join the dataframes
maryland_childcare_counties <- maryland_childcare_counties |> 
  inner_join(md, join_by(GEOID == GEOID))
# Calculate the rates for children per facility
maryland_childcare_counties <- maryland_childcare_counties |>
  mutate(child_care_rate= total/estimate.x) |>
  arrange(desc(child_care_rate))
```

##### **A3: Kent County**

------------------------------------------------------------------------

#### **Q4** Make a map of the results for question 3. Are there any areas of the state that are more lacking in childcare facilities?

```{r}
ggplot() +
  geom_sf(data=maryland_childcare_counties, aes(fill=child_care_rate)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log")

```

##### **A4**: Garrett, Worcester, and Anne Arrundel are lacking facilities

------------------------------------------------------------------------

#### **Q5** Now let's explore if we can find patterns in under-served areas, particularly when it comes to median income of the population. First, join the childcare data median income data from the Census. Then, create a scatterplot and perform a cor.test(), and answer: is there is any correlation between income and rate of children per facilities? What makes you conclude that?

```{r}
# Again, grab data from the ACS. The ACS variable for median income is B19013_001. We want 2022 data again.
md2 <- get_acs(geography = "county",
              variables = c(median_income = "B19013_001"),
              state = "MD",
              year = 2022)
```

```{r}
# Rename the column name 'estimate', which has our income data, with something more useful, like 'median_income'
md2 <- md2|>
  rename(median_income = estimate)
```

```{r}
# Join the dataframes
maryland_childcare_counties <- maryland_childcare_counties |> 
  inner_join(md2, join_by(GEOID == GEOID))
```

```{r}
# Create a scatterplot with median income and rate of children per facility
maryland_childcare_counties |>
  ggplot() +
  geom_point(aes(x=child_care_rate,y=median_income)) +
  geom_smooth(aes(x=child_care_rate,y=median_income), method="lm")
```

```{r}
# Create a correlation test.
cor.test(maryland_childcare_counties$child_care_rate, maryland_childcare_counties$median_income)
```

##### A5: There is a really small correlation

------------------------------------------------------------------------

#### **Q6** Finally, let's do some text analysis. We another dataset, with inspection reports from informal child care providers in the state. This dataset has an informative column of initial findings: a column that describes issues found in these providers, if any. Your goal is to create bigrams for this column and find common issues that are mentioned in these findings. And then answer: what kind of issues listed there you think are interesting? They don't have the most frequent.

```{r}
# Read the reports data
reports <- read_csv('data/childcare-informal-provider-inspections.csv') |> clean_names()
unique_words <- reports |> select(text) |>
  unnest_tokens(word, text)
View(unique_words)
```

```{r}
# Create a stopword dictionary (feel free to include more words)
data("stop_words") 
  stop_words <- stop_words |>
  add_row(word = 'complaint') |> 
  add_row(word = "H&S") |> 
  add_row(word = "standards") 
  
```

```{r}
# Unnest tokens into bigrams, removing stop words and repasting bigrams, list bigrams
reports |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  mutate(bigram = paste(word1, word2, sep=" ")) |>
  group_by(bigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

##### A6:
