---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, "As police struggle to solve homicides, Baltimore residents see an ‘open season for killing," written by Wesley Lowery, Steven Rich, and Salwan Georges. Link: <https://www.washingtonpost.com/investigations/as-police-struggle-to-solve-homicides-baltimore-residents-see-an-open-season-for-killing/2018/12/26/7ee561e4-fb24-11e8-8c9a-860ce2a8148f_story.html>

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
wbb_2022 <- read_csv("data/wbb_rosters_2022_23.csv") |> clean_names() 

wbb_2023 <-read_csv("data/wbb_rosters_2023_24.csv")|> clean_names() 

wbb_2021 <-read_csv("data/wbb_rosters_2021_22.csv")|> clean_names()

wbb_2020 <- read_csv("data/wbb_rosters_2020_21.csv")|> clean_names()

wbb_crosswalk <- read_csv("data/wbb_rosters23_crosswalk.csv") |> clean_names()


```

## Sentences to Engineer

### Sentence 1

-   **Sentence text**: What is the average height of players of each year
-   **Analysis summary**: 2020 has the highest average with 81 inches. Each year had a group of cities tied for the tallest average. A problem we had was we couldn't load 2021 because there was no total inches column. In 2020, the tallest height 6.75 feet and in 2022 and 2023 the tallest height was 6.583 ft.

```{r}
wbb_2020_height_per_town <- wbb_2020 |>
  group_by(hometown_clean) |> summarize(Avg_height_per_town = mean (total_inches)) |> arrange(desc(Avg_height_per_town)) |>
  mutate(feet = Avg_height_per_town/12)


wbb_2022_height_per_town <- wbb_2022 |>
   group_by(hometown_clean) |> mutate(total_inches) |> summarize(Avg_height_per_town = mean (total_inches)) |> arrange(desc(Avg_height_per_town)) |>
  mutate(feet = Avg_height_per_town/12)

wbb_2023_height_per_town <- wbb_2023 |>
  group_by(hometown_clean) |> mutate(total_inches) |> summarize(Avg_height_per_town = mean (total_inches)) |> arrange(desc(Avg_height_per_town)) |>
  mutate(feet = Avg_height_per_town/12)


```

### Sentence 2

-   **Sentence text**: How many people play in each position from 2020-2023
-   **Analysis summary**: The guard was the highest position for all 4 year with 7,670 in 2020, with 8,109 in 2021, with 8,305 in 2022 and with 8,282 in 2023.

```{r}
wbb_positions_2020 <- wbb_2020 |>
  group_by(position_clean) |>
  summarize(count = n())

wbb_positions_2021 <- wbb_2021 |>
  group_by(position_clean) |>
  summarize(count = n())

wbb_positions_2022 <- wbb_2022 |>
  group_by(position_clean) |>
  summarize(count = n())

wbb_positions_2023 <- wbb_2023 |>
  group_by(position_clean) |>
  summarize(count = n())

```

### Sentence 3

-   **Sentence text**: In the years 2022 and 2023, how many teams had more than 1 redshirt on their team? What team had the most, and where are they from
-   **Analysis summary**: In 2022 Grand Valley State in Michigan had the most red shirts with 12 and in 2023 Biola in California had the most red shirts 10.

```{r}
redshirt_22 <-wbb_2022 |>
  group_by(team, team_state) |>
  summarise(redshirt_numbers = sum(redshirt)) |> arrange(desc (redshirt_numbers)) |>
  filter(redshirt_numbers > 1)

redshirt_23 <- wbb_2023 |>
  group_by(team, team_state) |>
  summarise(redshirt_count = sum(redshirt)) |> arrange(desc (redshirt_count)) |>
  filter(redshirt_count > 1)



```

### Sentence 4

-   **Sentence text**: What state producers the most WNBA players

-   **Analysis summary**:

```{r}
wbb_states_2020 <- wbb_2020 |>
  group_by(homestate) |>
  summarize(count = n()) |>
  arrange(desc(count))

wbb_states_2021 <- wbb_2021 |>
  group_by(homestate) |>
  summarize(count = n()) |>
  arrange(desc(count))

wbb_states_2022 <- wbb_2022 |>
  group_by(homestate) |>
  summarize(count = n()) |>
  arrange(desc(count))

wbb_states_2023 <- wbb_2023 |>
  group_by(homestate) |>
  summarize(count = n()) |>
  arrange(desc(count))


```

### Sentence 5

-   **Sentence text**:

-   Analysis:

```{r}


```

-30-
