---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

murder_data <- read_csv("data/homicide-data.csv")
murder_data <- read_csv("data/homicide-data copy.csv")
```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: City police made an arrest in 41 percent of homicides in 2014
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]
    -   When you filter for just Baltimore and 2014, the original number of murders is 211. Then when you filter for closed by arrest you get 86. 86/211= .407 which is roughly 41%.

```{r}
# Put code to reverse engineer sentence here
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Baltimore")) |>
  filter (str_detect(disposition, "Closed by arrest"))
# Display results of code below this codeblock


```

### Sentence 2

-   **Sentence text**: Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest."
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
murder_data |>
 filter(str_detect(reported_date, "2015") | str_detect(reported_date, "2016") | str_detect(reported_date, "2017"))|>
  filter(str_detect(city, "Baltimore")) |>
  filter (str_detect(disposition, "Closed by arrest"))
# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: In Chicago, the homicide arrest rate has dropped 21 percentage points, in Boston it has dropped 12 points and in St. Louis it is down 9.
-   **Analysis summary**: 125/418= 29.9% , 68/654= 8.5% which gives Chicago a nearly 21% drop. 65/160= 40.6% , 64/202= 31.6% which gives St. Louis a 9% drop. 29/55= 52% , 22/54= 40% which gives Boston a 12% drop

```{r}
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Chicago")) 
#418
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Chicago")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#125
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Chicago")) 
#654
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Chicago")) |>
#68
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "St. Louis")) 
#160
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "St. Louis")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#65
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "St. Louis")) 
#202
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "St. Louis")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#64

murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Boston")) 
#55
murder_data |>
  filter(str_detect(reported_date, "2014")) |>
  filter(str_detect(city, "Boston")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#29
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Boston")) 
#54
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Boston")) |>
  filter (str_detect(disposition, "Closed by arrest"))
#22


# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**:

    Since 2015, the arrest rate hasn’t topped 30 percent in any year. (Baltimore)

-   **Analysis summary**:

    2017- 27.3% 2016 2016- 22.5% 2015- 25.4%

```{r}
# 2015 homicides 342
murder_data |>
  filter(str_detect(reported_date, "2015")) |>
  filter(str_detect(city, "Baltimore")) 

# 2015 closed 87
murder_data |>
  filter(str_detect(reported_date, "2015")) |>
  filter(str_detect(city, "Baltimore")) |>
  filter(str_detect(disposition, "Closed by arrest"))

#2016 homicides 320 
murder_data |>
  filter(str_detect(reported_date, "2016")) |>
  filter(str_detect(city, "Baltimore")) 

#2016 closed 72
murder_data |>
  filter(str_detect(reported_date, "2016")) |>
  filter(str_detect(city, "Baltimore")) |>
  filter(str_detect(disposition, "Closed by arrest"))

#2017 homicides 340
murder_data |>
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Baltimore")) 
#2017 closed 93
murder_data |> 
  filter(str_detect(reported_date, "2017")) |>
  filter(str_detect(city, "Baltimore")) |>
  filter(str_detect(disposition, "Closed by arrest"))
```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
